---
title: "pooled analysis of PK data fro Indometh"
author: "Andrew Carvalho"
date: '2024-03-15'
output: html_document
---
## Packages
```{r, echo=FALSE}
knitr::opts_chunk$set(comment='.', message=FALSE, fig.path="figures/indo-", 
                      cache=TRUE, autodep=TRUE)
options(mrgsolve.soloc="build", dplyr.summarise.inform = FALSE)
```

```{r}
library(tidyverse)
theme_set(theme_bw())
library(mrgsolve)
```



## load indamethacin dataset
```{r}
data("Indometh")
```

#take a peak


```{r}
head(Indometh)

count(Indometh, Subject)

ggplot(Indometh, aes(time, conc, group = Subject)) +
  geom_point() +
  scale_y_continuous(trans ="log", breaks = 10^seq(-4,4))
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


This is individual-level data, but we are going to do naive pooled 
analysis. 
# Assemble Data
```{r}
data <- readRDS("C:/Users/AndrewCarvalho/OneDrive - Apellis Pharmaceuticals, Inc/Documents/R/r-pharma-pkpd-2020-master/content/data/indometh.RDS")

head(data)
```
#load a pk model

```{r}
mod <- modlib("PK1")

param(mod)
```
# Now let's estimate some params

```{r}
theta <- log(c(CL = 1, V= 100))
```

```{r}
names(theta)
```
#Create an objective function function

- For starters, just do OLS estimation
- Note that we _need_ to name the parameters (`p`)
    - Parameter updates require names in `mrgsolve`
    - Generally, don't expect `p` to retain any names that you might
    pass in through the initial estimates
- We also pass in the `data` and the dependent variable (`dv`)

```{r}
obj <- function(p, theta, data, dv ="conc", pred = FALSE){
  
  names(p) <- names(theta)
  
  p <- lapply(p,exp)
  
  mod <- param(mod, p)
  
  out <- mrgsim_q(mod, data, output="df")
  
  if(pred) return(out)
  
  sqr <- (out[["CP"]] - data[[dv]])^2
  
  sum(sqr, na.rm=TRUE)
}
```
# Fit with one-compartment model

- First generate some initial estimates
- These _need_ to be named in a way that is consistent with
the model we are using
- I usually run a test with the objective function function to 
make sure the logic works out
```{r}
obj(theta,theta,data)
```
```{r}
fit <- optim(par = theta, fn=obj, theta = theta, data=data)
head (fit)
```

- And generate some predictions based on the final estimates
```{r}
pred <- obj(fit$par, theta, data, pred = TRUE)

data$pred <- pred$CP

head(data)
```
#make aplot of the output you just generated

```{r}
ggplot(data = data) +
  geom_point(aes(time,conc)) +
  scale_y_log10() +
  geom_line(aes(time,pred), col = "forest green", lwd =1.0)
```
# Fit doesnt look great so try fitting inometh data woth a 2cmpt model

```{r}
mod <- modlib("pk2")
```
```{r}
param(mod)

see(mod)
```
#set the initial estimates for the 2-comp model

```{r}
theta <- log(c(CL = 2, V2 = 50, Q = 10, V3 = 50))

```
# Fit

```{r}
fit <- optim(par = theta, fn=obj, theta = theta, data=data)

head(fit)
```

# now generate predictions on the fitted "final" estimates

```{r}
pred <- obj(fit$par, theta, data, pred = TRUE)

data$pred <- pred$CP


  ggplot(data = data) +
  geom_point(aes(time,conc)) +
  scale_y_log10() +
  geom_line(aes(time,pred), col = "forest green", lwd =1.0)
  
  head(data)
  
```


```{r}
 mod %>% ev(amt = 25, cmt =2) %>% 
  mrgsim() %>% plot()

```

```{r}

#change model parmaters from optimization above
mod <- param(mod, CL = 7.905, V2 = 8.9, Q = 10, V3 = 19.527)
# then view updated model 
param(mod)

mod %>% ev(amt = 25, cmt =2) %>% 
  mrgsim() %>% plot()
```


```{r}
#  The piped code I started with probably works nicely for some people, but it’s not my preferred way to do this. The way I think of these things, a simulation has three main inputs (model object, event schedule, simulation times), and I find the code easier to read when these three inputs are passed as three separate arguments. this would normally be the project folder


#dir <- fs::path_package("mrgsolve", "models")

# define model, events, and times

events <- ev_expand(amt = 100, ii = 24, addl = 9, cmt= 2)
times  <- tgrid(start = 0, end = 480, delta = 0.1)
```
```{r}
tibble::as_tibble(events)
```


```{r}
# run simulation
out <- mrgsim_d(x = mod, data = events, tgrid = times)

plot(out)
  
```
## do a ggplot of simlation in stead of the plot() function
The plot method allows you to customise the plot in a fairly flexible way, but there are some limits to what you can do with this approach. It’s not a big drawback though. If additional customisation is needed it’s pretty easy to convert the output to a tibble and then using ggplot2 to create the specific visualisation you want:


```{r}
# Adjust theme to make pretty plot
#theme_set(theme_bw(base_size=25))

library(ggplot2)

out %>% 
  tibble::as_tibble() %>% 
  tidyr::pivot_longer(
    cols = c(EV, CENT, CP), 
    names_to = "variable", 
    values_to = "value"
  ) %>%  
  dplyr::mutate(
    variable = dplyr::case_when(
      variable == "EV" ~ "Gut amount",
      variable == "CENT" ~ "Central amount",
      variable == "CP" ~ "Central concentration"
    )
  ) %>% 
  ggplot(aes(time, value)) + 
  geom_line() +
  facet_wrap(~ variable, scales = "free_y") + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) + 
  labs(
    title = "Drug amounts and concentrations over time",
    x = "Time (hours)",
    y = "mg or mg/L"
  )
   

```

```{r}
# define model, events, and times


events <- ev_expand(amt = 25, cmt =2 )
times  <- tgrid(start = 0, end = 8, delta = 0.1)
```



```{r}
out <- mrgsim_d(x = mod, data = events, tgrid = times)
out %>% 
  tibble::as_tibble() 
```
```{r}
data(Indometh)

head(Indometh)

count(Indometh, Subject)

```
#try to combine in sim and observed data in GGPlot



```{r}
library(ggplot2)
out %>% 
  tibble::as_tibble() %>% 
  
ggplot  +
  geom_point(data = tibble, mapping = aes( x = time, y = Plasma_mL), color = "Light Slate Blue")  + 
  
  geom_point(data = Indometh, mapping = aes( x= time, y = conc), color = "dark Orange") +
  #geom_line() +
  #facet_wrap(~ variable, scales = "free_y") + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) + 
  labs(
    title = "Drug amounts and concentrations over time",
    x = "Time (hours)",
    y = "mg or mg/L"
  )

```


```{r}

# now try to model theophyline PK data!!!
head(Theoph) 
```
```{r}

count(Theoph, Subject)

ggplot(data = Theoph, mapping = aes(x= Time, y= conc))+
         geom_smooth() +
  geom_point()
```
#load a model

```{r}
mod <- mread("popPK1cmpt.cpp", quiet = TRUE)
param(mod)

```
# define model, events, and times

```{r}
events <- ev_expand(amt = 4*70, ii = 24, addl = 0, ID = 1:12)
times  <- tgrid(start = 0, end = 24, delta = 0.5)
```

#Simlate events
```{r}
out <- mrgsim_d(x = mod, data = events, tgrid = times)
out %>% 
  tibble::as_tibble() 
```
```{r}
#change model parmaters from optimization above
mod <- param(mod, TVKA = 0.5, TVCL = 2, TVV = 24)
# then view updated model 
param(mod)
```
```{r}
out <- mrgsim_d(x = mod, data = events, tgrid = times) 
plot(out)
```
```{r}
library(ggplot2)
out %>% 
  tibble::as_tibble() %>% 
  
ggplot  +
  geom_point(data = tibble, mapping = aes( x = time, y = CP), color = "Light Slate Blue")  + 
  
  geom_point(data = Theoph, mapping = aes( x= Time, y = conc), color = "dark Orange") +
  #geom_line() +
  #facet_wrap(~ variable, scales = "Free_Y") + 
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) + 
  labs(
    title = "Drug amounts and concentrations over time",
    x = "Time (hours)",
    y = "mg or mg/L"
  )
```

